OA Share

### 2021-01-17 analysis/008_subject_exploration.Rmd
improve issn matching

```{r}
scopus_norm <- scopus_19 %>%
  gather(print_issn, e_issn, key = "issn_tpye", value = "issn") %>%
  # trailing zero's missing in Excel spreadsheet
  mutate(
    issn = ifelse(nchar(issn) == 5, paste0("000", issn), issn),
    issn = ifelse(nchar(issn) == 6, paste0("00", issn), issn),
    issn = ifelse(nchar(issn) == 7, paste0("0", issn), issn)) %>%
  # missing hyphen
  mutate(issn = map_chr(issn, function(x) paste(c(substr(x, 1, 4), substr(x, 5, 8)), collapse = "-")))
```

@@

### Correlation test

```{r}
hybrid_subject_19 <- hybrid_subject %>% 
  mutate(prop = oa /articles) %>%
  filter(issued_year == "2019", !is.na(percentile)
         #, prop > 0.01 & prop < 0.15) 
         )%>%
  distinct(issn, prop, percentile, snip, sjr)
```

spearman

```{r}
cor.test(hybrid_subject_19$prop, hybrid_subject_19$snip, method = "spearman")
```

plot
```{r}
ggplot(hybrid_subject_19, aes(prop, snip)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
  coord_cartesian(xlim = c(0., 0.7), ylim = c(0.5, 6)) +
  labs(x = "OA Uptake Rate", y = "SNIP") +
  theme_cowplot()
```

- cell system, global food security

```{r}
hybrid_subject %>%
  mutate(prop = oa /articles) %>%
  filter(issued_year == "2019", !is.na(percentile)) %>% 
  ggplot(aes(prop, snip)) + geom_point(alpha = .3) + geom_smooth(method = "lm") +
  facet_wrap(~top_level_code, scales = "free") +
  theme_minimal()
```

## OA proportion per subject

```{r}
# sort by median per asjc subject group
subject_all_years <- hybrid_subject %>% 
  ungroup() %>% 
  group_by(issn, description, subject_area) %>% 
  summarise(n_oa = sum(oa),
            n = sum(articles)) %>%
  ungroup() %>%
  group_by(description) %>%
  filter(subject_area != "Multidisciplinary")

median_all <- subject_all_years %>% 
  ungroup() %>% 
  distinct(issn, .keep_all = TRUE) %>% 
  mutate(prop = n_oa /n) %>% 
  .$prop %>% 
  median()

ggplot(subject_all_years, aes(fct_reorder(description, n_oa / n, .fun = median), n_oa / n)) + 
  geom_boxplot(aes(color = subject_area)) +
  geom_jitter(alpha = .05, aes(color = subject_area), size = 1) +
  geom_hline(yintercept = median_all,  colour = "red", linetype ="dashed", size = 1) +
  scale_colour_manual(NULL, values = c("#7F3C8D", "#11A579", "#3969AC", "#F2B701")) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 5L),
    expand = expansion(mult = c(0, 0.05))
  ) +
  coord_flip(ylim=c(0,0.2)) +
  labs(x = NULL, y = "Five-Years OA Uptake Rate per Journal") +
  theme_minimal_vgrid(font_size = 10) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(ncol = 2))
```

- Comptes Rendus https://www.library.illinois.edu/mtx/2020/05/01/comptes-rendus-de-lacademie-des-sciences-important-publishing-and-access-changes/

### by funder

```{r}
jn_subjects <- read_csv(here::here("data", "jn_subjects.csv"))

hybrid_volume_subjects <- hybrid_volume %>%
  group_by(oa_sponsor_type, articles, issued_year, issn) %>% 
  summarise(n_oa = n_distinct(doi)) %>% 
  left_join(jn_subjects, by = "issn") %>%
  ungroup()

hybrid_volume_subjects %>%
 # filter(container_title == "Journal of the American College of Cardiology") %>%
  group_by(top_level, articles, issued_year) %>%
  summarise(n_oa = sum(n_oa)) %>%
  ungroup() %>%
  group_by(top_level) %>%
  summarise(n_oa = sum(n_oa),
            n = sum(articles), .groups = "drop") %>%
  mutate(prop = n_oa / n) %>%
  arrange(desc(prop))

# hybrid_volume %>%
#   left_join(jn_subjects, by = "issn") %>%
#   group_by(oa_sponsor_type, issn, issued_year, top_level, articles) %>%
#   summarise(n_oa = n_distinct(doi)) %>%
#   group_by(oa_sponsor_type, issued_year, top_level) %>%
#   summarise(n_oa = sum(n_oa),
#             n = sum(articles)) %>%
#   ungroup() %>%
#   group_by(top_level, oa_sponsor_type) %>%
#   summarise(n_oa = sum(n_oa, na.rm = TRUE),
#             n = sum(n, na.rm = TRUE)) %>%
#   mutate(prop = n_oa / n) %>% View()
# 
# 
# hybrid_volume %>%
#   left_join(jn_subjects, by = "issn") %>%
#   group_by(issn, issued_year, top_level, article) %>%
#   summarise(n_oa = n_distinct(doi)) %>%
#   group_by(issued_year, top_level) %>%
#   summarise(nn = sum(n_oa))
```

```{r}
hybrid_volume_jn <- hybrid_volume %>%
  group_by(issn) %>%
  summarise(n_oa = n_distinct(doi))

share_per_invoicing_type <- hybrid_volume %>%
  group_by(oa_sponsor_type, issn) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  left_join(hybrid_volume_jn, by = "issn") %>%
  mutate(prop = n_oa_type / n_oa) %>%
  left_join(jn_subjects, by = "issn") %>%
  ungroup()
share_per_invoicing_type %>%
  filter(oa_sponsor_type == c("FundingBody")) %>%
  ggplot(aes(fct_reorder(top_level, prop, .fun = median, .desc = TRUE), prop)) +
  geom_boxplot() +
  coord_flip()
```

```{r}
hybrid_jn_volume <- hybrid_oa_volume %>%
  group_by(issn) %>% 
  summarise(n_oa = sum(oa))

hybrid_df <- readr::read_csv(here::here("data", "hybrid_articles.csv"))
funder_shares <- hybrid_df %>%
   mutate(oa_sponsor_type = ifelse(!oa_sponsor_type %in% c("FundingBody", "Author", "ElsevierWaived"), "Other", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  group_by(oa_sponsor_type, issn) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  left_join(hybrid_jn_volume, by = "issn") %>%
  mutate(prop = n_oa_type / n_oa) %>%
  left_join(jn_subjects, by = "issn") %>%
  ungroup() %>%
  group_by(top_level, oa_sponsor_type) %>%
  summarise(oa_type = sum(n_oa_type)) %>%
  mutate(prop = oa_type / sum(oa_type))
# overall props
oa_sponsor_type_all <- hybrid_df %>% 
  #exclude multidisciplinary
  filter(issn != "2095-9273") %>%
  count(oa_sponsor_type) %>% 
  mutate(prop = n / sum(n))

funder_shares %>% 
  filter(oa_sponsor_type %in% c("Author", "Agreement"), top_level != "Multidisciplinary") %>%
  pivot_wider(id_cols = top_level, names_from = oa_sponsor_type, values_from = prop) %>%
  filter(!is.na(top_level)) %>%
  ggplot() +
   ggalt::geom_dumbbell(
    aes(y = reorder(top_level, Author), x = Agreement, xend = Author),
    colour = "grey20",
    colour_xend = "#EDAE49",
    colour_x = "#30638E",
    size_x = 1.5,
    alpha = 0.9,
    size_xend = 1.5,
    show.legend = TRUE
   ) +
  geom_vline(xintercept = oa_sponsor_type_all %>% filter(oa_sponsor_type == "Author") %>% .$prop,  colour = "#EDAE49", linetype ="dashed", size = 1) +
    geom_vline(xintercept = oa_sponsor_type_all %>% filter(oa_sponsor_type == "FundingBody") %>% .$prop,  colour = "#30638E", linetype ="dashed", size = 1) +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 5L),
    expand = expansion(mult = c(0, 0.05)),
    limits = c(0,1)
  ) +
  labs(y = NULL, x = "Percentage Invoicing Type") +
  theme_minimal_vgrid(font_size = 10)
```

issues:

- high share of elseveir waived articles due to ome issue published in Nuclear and Particle Physics Proceedings: "37th International Conference on High Energy Physics (ICHEP)
(ICHEP 2014)2-9 July 2014, Valencia, Spain" Do we want to exclude this journal?

table representation

```{r}
subject_all <- funder_shares %>%
  group_by(top_level) %>%
  summarise(oa = sum(oa_type))
funder_overview_subject <- left_join(funder_shares, subject_all, by = "top_level")
# gt
funder_overview_subject %>%
  pivot_wider(id_cols = c("top_level", "oa"), names_from = oa_sponsor_type, values_from = prop) %>%
  ungroup() %>% 
  # remove na and multidisciplinary 
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  relocate(Agreement, .after = Author) %>%
  arrange(desc(Author)) %>%
  gt::gt(rowname_col = "top_level") %>%
  gt::cols_label(oa = "OA Articles") %>%
  fmt_percent(
    columns = vars(Author, Agreement, `Fee Waived`, Other),
    decimals = 0) %>%
  cols_width(
    2:6 ~ px(80)) %>%
  cols_align(
    align = "right",
    columns = 2:6
  ) %>%
  cols_width(
    2 ~ px(120)
  ) %>%
    gt::cols_align(
    align = "left",
    columns = 1
  ) %>%
  fmt_number(
    columns = vars(oa),
    decimals = 0
  ) %>%
  data_color(
    columns =3:6,
    colors = scales::col_numeric(
      # custom defined values - notice that order matters!
      palette = c("white", "#fcde9c","#faa476","#f0746e","#e34f6f","#dc3977","#b9257a","#7c1d6f"),
      domain = NULL
    )
  ) %>%
  tab_spanner(
    label = md("**OA Sponsor Type**"),
    columns = 3:6
  ) %>%
    tab_style(
    style = cell_text(color = "black", weight = "bold"),
    locations = list(
      cells_column_labels(everything())
    )) %>%
   tab_options(
    column_labels.border.top.color = "white",
    column_labels.border.top.width = px(3),
    column_labels.border.bottom.color = "black",
    table_body.hlines.color = "white",
    table.border.bottom.color = "white",
    table.border.bottom.width = px(3)
      )
```

## fields per funder

```{r}
field_funder <- hybrid_df %>%
  filter(oa_sponsor_type == "FundingBody") %>% 
  mutate(oa_sponsor = ifelse(is.na(oa_sponsor_name), "Other", oa_sponsor_name)) %>%
   mutate(oa_sponsor = fct_infreq(oa_sponsor)) %>%
  mutate(oa_sponsor = fct_lump_n(oa_sponsor, 10)) %>%
  left_join(jn_subjects, by = "issn") %>%
  filter(!is.na(top_level), top_level != "Multidisciplinary") %>%
  group_by(oa_sponsor, top_level) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  mutate(prop = n_oa_type / sum(n_oa_type)) %>%
  mutate(oa_sponsor = fct_reorder(oa_sponsor, n_oa_type))
ggplot(field_funder, aes(x = oa_sponsor, 
                         y = reorder(top_level, desc(top_level)), 
                         colour = prop, size = n_oa_type)) + 
  geom_point() +
  scale_color_viridis_c(labels = scales::percent_format(accuracy = 5L),
                      option = "magma",
                      name = "Percentage per funder",
                      breaks =  scales::extended_breaks()) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 6), position = "top") +
  labs(x = NULL, y = NULL) +
  theme_minimal_grid(font_size = 10) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank()) +
  theme(legend.position = "top", legend.justification = "right") +
  theme(legend.direction = "horizontal", legend.box = "vertical") 
```

 correlation funder / oa share

```{r}
funder_jn_volume <- hybrid_volume %>%
  group_by(oa_sponsor_type, issn) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  left_join(hybrid_volume_jn, by = "issn") %>%
  mutate(prop = n_oa_type / n_oa) %>%
 # left_join(jn_subjects, by = "issn") %>%
  ungroup() %>%
  filter(oa_sponsor_type == "FundingBody")
hybrid_funder_shares <- hybrid_volume %>%
  group_by(issn) %>%
  summarise(all = sum(articles)) %>%
  inner_join(funder_jn_volume, by = "issn") %>%
  mutate(oa_prop = n_oa / all)

  ggplot(hybrid_funder_shares, aes(oa_prop, prop)) +
  geom_point(alpha = .3) + 
  geom_smooth(method = "lm") +
  theme_cowplot()
```

```{r}
cor.test(hybrid_funder_shares$prop, hybrid_funder_shares$oa_prop, method = "spearman")
```

oa sponsor name subject

```{r}
test_a <- hybrid_volume %>%
  filter(oa_sponsor_type == "FundingBody") %>%
  group_by(oa_sponsor_name, issn) %>%
  summarise(n_oa_type = n_distinct(doi)) %>%
  left_join(hybrid_volume_jn, by = "issn") %>%
  mutate(prop = n_oa_type / n_oa) %>%
  left_join(jn_subjects, by = "issn") %>%
  ungroup()  %>%
  group_by(top_level, oa_sponsor_name) %>%
  summarise(oa = sum(n_oa_type))
```

```{r}
tt <- test_a %>%
  pivot_wider(names_from = "oa_sponsor_name", values_from = "oa")
wilcox.test(tt$`Bill & Melinda Gates Foundation`, tt$VSNU, paired = TRUE)
```

>>>>

```{r}
scopus_norm <- scopus_19 %>%
  separate(print_issn, c("print_issn", "print_issn_extra")) %>% 
  gather(print_issn, e_issn, print_issn_extra, key = "issn_tpye", value = "issn") %>%
  # trailing zero's missing in Excel spreadsheet
  mutate(
    issn = ifelse(nchar(issn) == 5, paste0("000", issn), issn),
    issn = ifelse(nchar(issn) == 6, paste0("00", issn), issn),
    issn = ifelse(nchar(issn) == 7, paste0("0", issn), issn)) %>%
  # missing hyphen
  mutate(issn = map_chr(issn, function(x) paste(c(substr(x, 1, 4), substr(x, 5, 8)), collapse = "-")))
```

@@
